Reduced plasticity and variance in physiological rates for terrestrial ectotherm populations under climate change

### 2024-01-08 ms.qmd
Just fixing figures so that they are using negative and ditching arrows. Need to also fix up the posteriors so they are not using absolute. We can take absolute later

# % change in acclimation on log scale
  change_f <- abs(f_acclim - f_acute)*100; quantile(change_f, c(0.025, 0.975))
  change_m <- abs(m_acclim - m_acute)*100; quantile(change_m, c(0.025, 0.975))
  change_t <- abs(t_acclim - t_acute)*100; quantile(change_t, c(0.025, 0.975))
  avg_change <- (change_f + change_m + change_t) / 3; quantile(avg_change, c(0.025, 0.975)) # Could use weighted mean, but for now this seems fine.

@@

Overall, $lnRR_{Q_{10_{acclimation}}}$ and $lnRR_{Q_{10_{acute}}}$ differed by only `r mean(avg_change)`% across all habitats (95%CI: `r quantile(avg_change, c(0.025, 0.975))[1]` to `r quantile(avg_change, c(0.025, 0.975))[2]`%). Ectotherms in marine and freshwater environments showed partial compensation of physiological rates ([@fig-1]B) amounting to reduced $lnRR_{Q_{10_{acclimation}}}$ of `r mean(change_f)`% (95% CI: `r quantile(change_f, c(0.025, 0.975))[1]` to `r quantile(change_f, c(0.025, 0.975))[2]`) in freshwater and `r mean(change_m)`% (95% CI: `r quantile(change_m, c(0.025, 0.975))[1]` to `r quantile(change_m, c(0.025, 0.975))[2]`) in marine environments. In contrast, terrestrial ectotherms showed no acclimation (possibly even inverse acclimation) -- showing a `r mean(change_t)`% increase in $lnRR_{Q_{10_{acclimation}}}$ (95% CI: `r quantile(change_t, c(0.025, 0.975))[1]` to `r quantile(change_t, c(0.025, 0.975))[2]`, [@fig-1]B). Considering acute responses of animals acclimated to high temperatures are generally slightly elevated compared to cold acclimated animals (~7%; $\beta$ = `r model1.s1$b[3]`, 95% CI: `r model1.s1$ci.lb[3]` to `r model1.s1$ci.ub[3]`, p `r p_value(model1.s1$pval[3])`), acclimatization is not likely going to provide adaptive benefits under climate change.

@@

p1.2 <- orchaRd::orchard_plot(model1.2, mod = "habitat", group  = "record_num", at = list(type = c("acute", "acclim")), by = "type", weights = "prop", xlab = TeX("$lnRR_{Q_{10}}$"), angle = 45, legend.pos = "top.left", trunk.size = 8, branch.size = 2) + scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +
  theme(legend.direction = "vertical") +
  geom_segment(aes(x = 2.92, y = 2, xend = 3.07, yend = 2), 
                            arrow = arrow(angle = 90, ends = "both", length = unit(0.01, "npc"))) +
  geom_segment(aes(x = 1.92, y = 2, xend = 2.07, yend = 2), 
                            arrow = arrow(angle = 90, ends = "both", length = unit(0.01, "npc"))) +
  geom_segment(aes(x = 0.92, y = 2, xend = 1.07, yend = 2), 
                            arrow = arrow(angle = 90, ends = "both", length = unit(0.01, "npc"))) +
annotate("text", x = 1, y = 2.10, 
         label =TeX(paste0("$\\beta = $", round(model1.2$b[3], 2), "[", round(model1.2$ci.lb[3], 2)," to ", round(model1.2$ci.ub[3], 2), "] ")), hjust = 0, size = 6) + 
annotate("text", x = 1.01, y = 4.0, 
         label =paste0(round(mean(change_f), 2), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
  geom_segment(aes(x = 0.92, y = 3.9, xend = 1.07, yend = 3.9), 
                            arrow = arrow(angle = 45, ends = "first", length = unit(0.01, "npc"))) +
  annotate("text", x = 0.85, y = 3, label = "P < 0.001", hjust = 0, size = 6) + 
  annotate("text", x = 2, y = 2.10, 
         label =TeX(paste0("$\\beta = $", round(mean(m_acclim - m_acute), 2), "[", round(quantile(m_acclim - m_acute, 0.025), 2)," to ", round(quantile(m_acclim - m_acute, 0.975),2), "] ")), hjust = 0, size = 6) + 
annotate("text", x = 2.01, y = 4.0, 
         label =paste0(round(mean(change_m), 2), "%"), hjust = 0, size = 6) +
  geom_segment(aes(x = 1.92, y = 3.9, xend = 2.07, yend = 3.9), 
                            arrow = arrow(angle = 45, ends = "first", length = unit(0.01, "npc"))) +
  annotate("text", x = 1.85, y = 3, label = "P = 0.01", hjust = 0, size = 6) +
  annotate("text", x = 3, y = 2.10, 
         label =TeX(paste0("$\\beta = $", round(mean(t_acclim - t_acute), 2), "[", round(quantile(t_acclim - t_acute, 0.025), 2)," to ", round(quantile(t_acclim - t_acute, 0.975),2), "] ")), hjust = 0, size = 6) + 
annotate("text", x = 3.01, y = 4.0, 
         label =paste0(round(mean(change_t), 2), "%"), hjust = 0, size = 6) +
  geom_segment(aes(x = 2.92, y = 3.9, xend = 3.07, yend = 3.9), 
                            arrow = arrow(angle = 45, ends = "last", length = unit(0.01, "npc"))) +
  annotate("text", x = 2.85, y = 3, label = "P = 0.36", hjust = 0, size = 6) + scale_fill_manual(values = c(col)) + ylim(-0.8, 5.05) +  annotate("text", label= paste0("k = ", k_lnrr$k, " (", k_lnrr$sdy, ")"), x = k_lnrr$moderator, y = 5)

@@

  p2 <- orchard_plot(model2.2, xlab = TeX("$lnCVR_{Q_{10}}$"), mod = "habitat", group  = "record_num", weights = "prop", angle = 45, trunk.size = 8, branch.size = 2) +
              scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) + annotate("text", x = 1.3, y = -1.8, 
           label =TeX(paste0("$\\mu = $", round(mean(f_cvr_mean), 2), "[", round(quantile(f_cvr_mean, 0.025), 2)," to ", round(quantile(f_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 1.15, y = -1.5, label = paste0("P = ", round(pmcmc(f_cvr_mean), digit = 2)), hjust = 0, size = 6) + 
  annotate("text", x = 1.15, y = 1, 
           label =paste0(round(mean(exp(f_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + 
  annotate("text", x = 2.3, y = -1.8, 
           label =TeX(paste0("$\\mu = $", round(mean(m_cvr_mean), 2), "[", round(quantile(m_cvr_mean, 0.025), 2)," to ", round(quantile(m_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 2.15, y = -1.5, label = paste0("P = ", round(pmcmc(m_cvr_mean), digit = 2)), hjust = 0, size = 6) + 
  annotate("text", x = 2.15, y = 1, 
           label =paste0(round(mean(exp(m_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + annotate("text", x = 3.4, y = -1.8, 
           label =TeX(paste0("$\\mu = $", round(mean(t_cvr_mean), 2), "[", round(quantile(t_cvr_mean, 0.025), 2)," to ", round(quantile(t_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 3.25, y = 1, 
           label =paste0(round(mean(exp(t_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
    annotate("text", x = 3.25, y = -1.5, label = paste0("P = ", round(pmcmc(t_cvr_mean), digit = 3)), hjust = 0, size = 6) + scale_fill_manual(values = c(col)) + ylim(-2, 1.8) + 
    annotate("text", label= paste0("k = ", k_cvr$k, " (", k_cvr$sdy, ")"), x = k_cvr$moderator, y = 2.4)

  p2 <- p2 + theme_classic() + theme(axis.title  = element_text(size = 24),
          axis.text = element_text(size = 15))

@@

  #(phylo) / (p1.2 / p2) + plot_annotation(tag_levels = "A", tag_suffix = ")")

>>>>

# % change in acclimation on log scale
  change_f <- (f_acclim - f_acute)*100; quantile(change_f, c(0.025, 0.975))
  change_m <- (m_acclim - m_acute)*100; quantile(change_m, c(0.025, 0.975))
  change_t <- (t_acclim - t_acute)*100; quantile(change_t, c(0.025, 0.975))
  avg_change <- (change_f + change_m + change_t) / 3; quantile(avg_change, c(0.025, 0.975)) # Could use weighted mean, but for now this seems fine.

@@

Overall, $lnRR_{Q_{10_{acclimation}}}$ and $lnRR_{Q_{10_{acute}}}$ differed by only `r mean(avg_change)`% across all habitats (95%CI: `r quantile(avg_change, c(0.025, 0.975))[1]` to `r quantile(avg_change, c(0.025, 0.975))[2]`%). Ectotherms in marine and freshwater environments showed partial compensation of physiological rates ([@fig-1]B) amounting to reduced $lnRR_{Q_{10_{acclimation}}}$ of `r mean(abs(change_f))`% (95% CI: `r quantile(abs(change_f), c(0.025, 0.975))[1]` to `r quantile(abs(change_f), c(0.025, 0.975))[2]`) in freshwater and `r mean(abs(change_m))`% (95% CI: `r quantile(mean(abs(change_m), c(0.025, 0.975))[1]` to `r quantile(mean(abs(change_m), c(0.025, 0.975))[2]`) in marine environments. In contrast, terrestrial ectotherms showed no acclimation (possibly even inverse acclimation) -- showing a `r mean(abs(change_t))`% increase in $lnRR_{Q_{10_{acclimation}}}$ (95% CI: `r quantile(abs(change_t), c(0.025, 0.975))[1]` to `r quantile(abs(change_t), c(0.025, 0.975))[2]`, [@fig-1]B). Considering acute responses of animals acclimated to high temperatures are generally slightly elevated compared to cold acclimated animals (~7%; $\beta$ = `r model1.s1$b[3]`, 95% CI: `r model1.s1$ci.lb[3]` to `r model1.s1$ci.ub[3]`, p `r p_value(model1.s1$pval[3])`), acclimatization is not likely going to provide adaptive benefits under climate change.

@@

p1.2 <- orchaRd::orchard_plot(model1.2, mod = "habitat", group  = "record_num", at = list(type = c("acute", "acclim")), by = "type", weights = "prop", xlab = TeX("$lnRR_{Q_{10}}$"), angle = 45, legend.pos = "top.left", trunk.size = 8, branch.size = 2) + scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +
  theme(legend.direction = "vertical") +
  geom_segment(aes(x = 2.92, y = 2, xend = 3.07, yend = 2), 
                            arrow = arrow(angle = 90, ends = "both", length = unit(0.01, "npc"))) +
  geom_segment(aes(x = 1.92, y = 2, xend = 2.07, yend = 2), 
                            arrow = arrow(angle = 90, ends = "both", length = unit(0.01, "npc"))) +
  geom_segment(aes(x = 0.92, y = 2, xend = 1.07, yend = 2), 
                            arrow = arrow(angle = 90, ends = "both", length = unit(0.01, "npc"))) +
annotate("text", x = 1, y = 2.10, 
         label =TeX(paste0("$\\beta = $", round(model1.2$b[3], 2), "[", round(model1.2$ci.lb[3], 2)," to ", round(model1.2$ci.ub[3], 2), "] ")), hjust = 0, size = 6) + 
annotate("text", x = 1.01, y = 4.0, 
         label =paste0(round(mean(change_f), 2), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
  annotate("text", x = 0.85, y = 3, label = "P < 0.001", hjust = 0, size = 6) + 
  annotate("text", x = 2, y = 2.10, 
         label =TeX(paste0("$\\beta = $", round(mean(m_acclim - m_acute), 2), "[", round(quantile(m_acclim - m_acute, 0.025), 2)," to ", round(quantile(m_acclim - m_acute, 0.975),2), "] ")), hjust = 0, size = 6) + 
annotate("text", x = 2.01, y = 4.0, 
         label =paste0(round(mean(change_m), 2), "%"), hjust = 0, size = 6) +
  annotate("text", x = 1.85, y = 3, label = "P = 0.01", hjust = 0, size = 6) +
  annotate("text", x = 3, y = 2.10, 
         label =TeX(paste0("$\\beta = $", round(mean(t_acclim - t_acute), 2), "[", round(quantile(t_acclim - t_acute, 0.025), 2)," to ", round(quantile(t_acclim - t_acute, 0.975),2), "] ")), hjust = 0, size = 6) + 
annotate("text", x = 3.01, y = 4.0, 
         label =paste0(round(mean(change_t), 2), "%"), hjust = 0, size = 6) +
  annotate("text", x = 2.85, y = 3, label = "P = 0.36", hjust = 0, size = 6) + scale_fill_manual(values = c(col)) + ylim(-0.8, 5.05) +  annotate("text", label= paste0("k = ", k_lnrr$k, " (", k_lnrr$sdy, ")"), x = k_lnrr$moderator, y = 5)

@@

  p2 <- orchard_plot(model2.2, xlab = TeX("$lnCVR_{Q_{10}}$"), mod = "habitat", group  = "record_num", weights = "prop", angle = 45, trunk.size = 8, branch.size = 2) +
              scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) + annotate("text", x = 1.3, y = -1.8, 
           label =TeX(paste0("$\\mu = $", round(mean(f_cvr_mean), 2), "[", round(quantile(f_cvr_mean, 0.025), 2)," to ", round(quantile(f_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 1.15, y = -1.5, label = paste0("P = ", round(pmcmc(f_cvr_mean), digit = 2)), hjust = 0, size = 6) + 
  annotate("text", x = 1.15, y = 1, 
           label =paste0(round(mean(exp(f_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + 
  annotate("text", x = 2.3, y = -1.8, 
           label =TeX(paste0("$\\mu = $", round(mean(m_cvr_mean), 2), "[", round(quantile(m_cvr_mean, 0.025), 2)," to ", round(quantile(m_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 2.15, y = -1.5, label = paste0("P = ", round(pmcmc(m_cvr_mean), digit = 2)), hjust = 0, size = 6) +
  annotate("text", x = 2.15, y = 1, 
           label =paste0(round(mean(exp(m_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + annotate("text", x = 3.4, y = -1.8, 
           label =TeX(paste0("$\\mu = $", round(mean(t_cvr_mean), 2), "[", round(quantile(t_cvr_mean, 0.025), 2)," to ", round(quantile(t_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = 3.25, y = 1, 
           label =paste0(round(mean(exp(t_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
    annotate("text", x = 3.25, y = -1.5, label = paste0("P = ", round(pmcmc(t_cvr_mean), digit = 3)), hjust = 0, size = 6) + scale_fill_manual(values = c(col)) + ylim(-2, 2) + 
    annotate("text", label= paste0("k = ", k_cvr$k, " (", k_cvr$sdy, ")"), x = c(1.15, 2.15, 3.25), y = 1.8)

  p2 <- p2 + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 15),
                                       legend.title = element_text(size = 15),
                                        legend.text = element_text(size = 15),)

@@
